No Replication Without Process Documentation:  An Assessment of Management

### 2020-04-10 R/MSE-manuscript.Rmd
minor text edits for spelling and grammar

---
title: 'No Replication Without Process Documentation:  An Assessment of Management
  Strategy Evaluations'
author:
- affiliation: School for Marine Science and Technology, University of Massachusetts
    - Dartmouth
  name: Jonathan W. Cummings, Amanda Hart, Gavin Fay
output:
  bookdown::html_document2:
    number_sections: false
  bookdown::word_document2:
  bookdown::pdf_document2:
    number_sections: false
always_allow_html: yes
biblio-style: apalike
endnote: no
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
header-includes: \usepackage{hyperref}
keywords: pandoc, r markdown, knitr
bibliography: MSEreview.bib
abstract: The aim and vision for fisheries management that comes from management strategy evaluation and adaptive management is a noble and aspirational vision.  We review MSE documentation in the published literature to appraise the achievement of this vision, finding that reality doesn't fully reflect this vision. The steps taken to complete the management strategy evaluation process often are not documented. Scientists often filled the role of fishery stakeholders, seemingly selecting the objectives and alternatives to evaluate in most management strategy evaluations. Facilitators and decision analysts -- participants focused on the decision making process rather than making the decision -- were rarely participants in MSE processes.  Conservation, Economic, and Yield focused objectives were commonly considered in MSE processes, while broader social objectives were rarely considered.  The picture painted by MSE documentation is primarily one of scientist driven simulation studies rather than management driven decision processes, with limited process documentation occurring. We hope providing a repository for reviews of previous management strategy evaluations will aid and speed learning among practitioners of management strategy evaluation.
---

@@

### Objectives 
Our overall objective is to determine the extent that MSE achieves the aspirational aims outlined above, and whether MSE documentation supports learning within the MSE practitioner community.  Given that scientific publication is the primary means of communicating the results of MSE, we review the scientific literature to assess the extent to which the process and aims of MSE, and adaptive management are being achieved and supported by the scientific literature. To assess this we determine how MSEs utilize standard components of decision making processes. We use the structured decision making (SDM) process [@SDM] -- the decision making framework in which adaptive management occurs -- as our framework for evaluation.  Using this framework we review published MSEs to:

@@

## Reviewing MSE documentation 
We used the structured decision making framework [@conroy2013decision; @gregory2012structured] as our framework for evaluating management strategy evaluation documents (Figure \@ref(fig:SDMfigure)). The structured decision making processes is composed of the six process stages: problem definition, objective elicitation, alternative development, consequence prediction, trade-off analysis, and decision implementation.

@@

In the participant group we address how participation in the MSE process was structured and who participated. We report who lead the MSE process, who participated in it, and if the roles of participants were established and documented.  In many cases the process used to elicit objectives and alternatives was not explicit.  In these cases we used what information was provided to glean who participated in the production of those components of the MSE.  We record and report who those participants were, referring to them as subjective participants in the objectives or alternatives process.  In other cases where there was not enough information to glean who participated we report the participants as unknown.

@@

Climate change MSEs were more likely to explicitly document the trade-off analysis stage, equally likely to document the alternatives stages, but were less likely to document the other stages in the decision process.
```{r, CCstages}
freq.data_CC %>%
  mutate("Percent"=round(Percent,0)) %>% 
  select(Explicit, Percent, Number, -Analysis) %>% 
  kable(caption = "Prevalence of explicit documentation of stages in climate change MSE processes") %>%
  kable_styling()
```

@@

The MSE process may be aided by utilizing the techniques from the structured decision making approach to decision analysis, such as expanding the focus on problem framing at the outset of MSEs, eliciting fully representative sets of objectives, and utilizing multiple attribute utility theory to evaluate tradeoffs between objectives (Cummings, in prep). Utilizing the structured decision making approach also eases the documentation process by breaking decision making into more discrete stages. However, there may be practical considerations, such as time or budgetary constraints that limit how complex the stages of a MSE are, and the degree of documentation that occurs.  Where the expertise and budget is available, the use of facilitators and scribes may aid in addressing and documenting MSE processes.

>>>>

---
title: 'No Replication Without Process Documentation:  An Assessment of Management
  Strategy Evaluations'
author:
- affiliation: School for Marine Science and Technology, University of Massachusetts
    - Dartmouth
  name: Jonathan W. Cummings, Amanda Hart, Gavin Fay
output:
  bookdown::word_document2:
  bookdown::html_document2:
    number_sections: false
  bookdown::pdf_document2:
    number_sections: false
always_allow_html: yes
biblio-style: apalike
endnote: no
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
header-includes: \usepackage{hyperref}
keywords: pandoc, r markdown, knitr
bibliography: MSEreview.bib
abstract: The aim and vision for fisheries management that comes from management strategy evaluation and adaptive management is a noble and aspirational vision.  We review MSE documentation in the published literature to appraise the achievement of this vision, finding that reality doesn't fully reflect this vision. The steps taken to complete the management strategy evaluation process often are not documented. Scientists often filled the role of fishery stakeholders, seemingly selecting the objectives and alternatives to evaluate in most management strategy evaluations. Facilitators and decision analysts -- participants focused on the decision making process rather than making the decision -- were rarely participants in MSE processes.  Conservation, Economic, and Yield focused objectives were commonly considered in MSE processes, while broader social objectives were rarely considered.  The picture painted by MSE documentation is primarily one of scientist driven simulation studies rather than management driven decision processes, with limited process documentation occurring. We hope providing a repository for reviews of previous management strategy evaluations will aid and speed learning among practitioners of management strategy evaluation.
---

@@

### Objectives 
Our overall objective is to determine the extent that MSE achieves the aspirational aims outlined above, and whether MSE documentation supports learning within the MSE practitioner community.  Given that scientific publication is the primary means of communicating the results of MSE, we review the scientific literature to assess the extent to which the process and aims of MSE, and adaptive management are being achieved and supported by the scientific literature. To assess this, we determine how MSEs utilize standard components of decision making processes. We use the structured decision making (SDM) process [@SDM] -- the decision making framework in which adaptive management occurs -- as our framework for evaluation.  Using this framework, we review published MSEs to:

@@

## Reviewing MSE documentation 
We used the structured decision making framework [@conroy2013decision; @gregory2012structured] as our framework for evaluating management strategy evaluation documents (Figure \@ref(fig:SDMfigure)). The structured decision making process is composed of the six process stages: problem definition, objective elicitation, alternative development, consequence prediction, trade-off analysis, and decision implementation.

@@

In the participant group we address how participation in the MSE process was structured and who participated. We report who lead the MSE process, who participated in it, and if the roles of participants were established and documented.  In many cases the process used to elicit objectives and alternatives was not explicit.  In these cases, we used what information was provided to glean who participated in the production of those components of the MSE.  We record and report who those participants were, referring to them as subjective participants in the objectives or alternatives process.  In other cases where there was not enough information to glean who participated we report the participants as unknown.

@@

Climate change MSEs were more likely to explicitly document the trade-off analysis stage, equally likely to document the alternatives stages, but were less likely to document the other stages in the decision process (Table \@ref(tab:CCstages)).
```{r, CCstages}
freq.data_CC %>%
  mutate("Percent"=round(Percent,0)) %>% 
  select(Explicit, Percent, Number, -Analysis) %>% 
  kable(caption = "Prevalence of explicit documentation of stages in climate change MSE processes") %>%
  kable_styling()
```

@@

The MSE process may be aided by utilizing the techniques from the structured decision making approach to decision analysis, such as expanding the focus on problem framing at the outset of MSEs, eliciting fully representative sets of objectives, and utilizing multiple attribute utility theory to evaluate tradeoffs between objectives (Cummings, in prep). Utilizing the structured decision making approach also eases the documentation process by breaking decision making into more discrete stages. However, there may be practical considerations, such as time or budgetary constraints that limit how complex the stages of a MSE are, and the degree of documentation that occurs.  Where the expertise and budget are available, the use of facilitators and scribes may aid in addressing and documenting MSE processes.
